Experience Report: Haskell in Computational Biology

### 2012-03-11 mrfy_experience_report.tex
tex file scrubbed clean of all latex errors

\authorinfo{Noah M. Daniels \and Andrew Gallant \and Norman Ramsey}
           {Department of Computer Science, Tufts University}
           {\{ndaniels, agallant, nr\}@cs.tufts.edu}

@@

@@

MRFy's search works with a sequence of \emph{generations}.
A~generation is a collection of scored \emph{placements}.
We~pass the
search function a \texttt{Scorer}, which is the result of partially
applying \texttt{score} to the model.
\smallverbatiminput{scoredecl}
A~search algorithm is characterized by four functions, which are
bundled into a record called \texttt{SearchStrategy}.
\smallverbatiminput{strategy}
In this record,
\begin{itemize}
\item
Function \texttt{gen0} produces the first generation of placements.
\item
Function \texttt{nextGen} produces a new generation of placements from
an existing generation.
Because scoring a placement is expensive, 
and some placements may persist from the old generation to the new
generation, \texttt{nextGen} works with \emph{scored} placements, not
bare placements.
\item
A~new generation produced by \texttt{nextGen} may not be any better
than the generation it replaces.
When it is first produced, the new generation is ``on trial.''
The \texttt{accept} function decides whether to keep a new generation
or to try \texttt{nextGen} again with a different seed.
The \texttt{accept} considers the score of the best
placement of the trial generation, along with the history of previous
best scores.
It~also considers the \texttt{Age} of the trial generation, which is
the total number of 
generations tried.
\item
Finally, the \texttt{quit} function decides when to stop the search.
It~too considers the history of scores and the age of the current
generation. 
\end{itemize}
This interface provides considerable leeway for experiments.
For~example, random-mutation hill climbing accepts only generations
with better scores than their predecessors,
while simulated annealing may accept a generation with a worse score.
A~genetic algorithm uses \texttt{nextGen} to
combine placements from the previous generation, while
random-mutation
hill climbing simply permutes a placement into a new one.
The interface \texttt{quit} makes it easy
to implement such termination criteria as a fixed number of generations
or a limit of going $k$~generations without achieving more than $E$
improvement in score.

@@

We hope you find the search code simple and straightforward:
\smallverbatiminput{search}
\emph{[MENTION AND QUANTIFY NOAH'S SUCCESSES IN ADDING NEW HEURISTICS.]}

@@

@@

\begingroup
\parfillskip=0pt
\advance\leftskip by 0pt plus 30em
\emph{We live to surf the Haskell wave, but~if you slide off the crest, you
drown.}
\par
\endgroup

>>>>

\authorinfo{Noah M. Daniels \and Andrew Gallant \and Norman Ramsey}
           {Department of Computer Science, Tufts University}
           {\char`\{ndaniels, agallant, nr\char`\}@cs.tufts.edu}

@@

\seclabel{hofs}
\seclabel{mrfy}

@@

MRFy's search works with a sequence of \emph{generations}.
A~generation is a collection of scored \emph{placements}.
We~pass the
search function a \texttt{Scorer}, which is the result of partially
applying \texttt{score} to the model.
\smallverbatiminput{scoredecl}
A~search algorithm is characterized by four functions, which are
bundled into a record called \texttt{SearchStrategy}.
\smallfuzzverbatiminput{2.3pt}{strategy}
In this record,
\begin{itemize}
\item
Function \texttt{gen0} produces the first generation of placements.
\item
Function \texttt{nextGen} produces a new generation of placements from
an existing generation.
Because scoring a placement is expensive, 
and some placements may persist from the old generation to the new
generation, \texttt{nextGen} works with \emph{scored} placements, not
bare placements.
\item
A~new generation produced by \texttt{nextGen} may not be any better
than the generation it replaces.
When it is first produced, the new generation is ``on trial.''
The \texttt{accept} function decides whether to keep a new generation
or to try \texttt{nextGen} again with a different seed.
The \texttt{accept} considers the score of the best
placement of the trial generation, along with the history of previous
best scores.
It~also considers the \texttt{Age} of the trial generation, which is
the total number of 
generations tried.
\item
Finally, the \texttt{quit} function decides when to stop the search.
It~too considers the history of scores and the age of the current
generation. 
\end{itemize}
This interface provides ample leeway for experiments.
For~example, random-mutation hill climbing accepts only generations
with better scores than their predecessors,
while simulated annealing may accept a generation with a worse score.
A~genetic algorithm uses \texttt{nextGen} to
combine placements from the previous generation, while
random-mutation
hill climbing simply permutes a placement into a new one.
The interface \texttt{quit} makes it easy
to implement such termination criteria as a fixed number of generations
or a limit of going $k$~generations without achieving more than $E$
improvement in score.

@@

We hope you find the search code simple and straightforward:
\smallfuzzverbatiminput{11pt}{search}
\emph{[MENTION AND QUANTIFY NOAH'S SUCCESSES IN ADDING NEW HEURISTICS.]}

@@

\seclabel{penumbra}

@@

\begingroup
\parfillskip=0pt
\advance\leftskip by 0pt plus 30em
\emph{You can live to surf the Haskell wave, but~if you slide off the crest, you
drown.}
\par
\endgroup
